---
title: "Exponential Distribution and Central Limit Theorem"
author: "Shirley Asangi"
fontsize: 12pt
output: pdf_document
---

## Synopsis
The Aim of this work is to investigate the exponential distribution and compare it with **Central Limit Theorem** to see if it follows the theorem or not.

## Simulation for Sampling Distriution of Exponential Distribution

```{r,fig.align="center",fig.cap="Density Plot for Distribution of Sample Means",fig.height=3}
nsim <- 1000
sample.dist <- vector(length=nsim)
nsamples <- 40
lambda <- 0.2
true.mean <- 1/lambda
true.variance <- (1/lambda)^2
variance <- vector(length=nsim)
for (i in 1:nsim){
  
  sample <- rexp(nsamples,rate=lambda)
  sample.dist[i] <- mean(sample)
  variance[i] <- var(sample)
}

library(ggplot2)
sample.mean <- mean(sample.dist)

variance.mean <- mean(variance)

data <- data.frame(x=sample.dist)
g <- ggplot(data, aes(x = x)) + geom_histogram(binwidth=.3, colour = "black", 
                                               aes(y = ..density..)) 
g <- g + geom_vline(xintercept=true.mean,color='red',size=2) + 
geom_vline(xintercept=sample.mean,color='blue',size=2) +
ggtitle("Distribution of Sample Means") + 
theme(plot.title = element_text(hjust = 0.5))
g
```

Blue Vertical Line shows the mean of the sample distribution and red line is the true mean of the exponential distribution i.e 1/lambda.The Sample Mean is a statistic trying to estimate true mean of distribution.The **Central Limit Theorem states that if you have a population with mean mu and standard deviation sigma and take sufficiently large random samples from the population with replacement , then the distribution of the sample means will be approximately normally distributed and its mean will be equal to true mean of distribution**.From above simulation plot this theory is confirmed as the mean of sample means is almost equal to true mean.
```{r}
print(paste0("The True Mean is ",true.mean))
print(paste0("The Mean of Sample Means is ",sample.mean))
```
\newpage
## Variance of samples and theoretical variance

```{r,fig.cap="Plot for variances of every sample",fig.align='center',fig.height=3}

sample.number = (1:length(variance))
variance.data <- data.frame(var=variance,sample.number=sample.number)
ggplot(data=variance.data) + geom_line(aes(y=var,x=sample.number)) + geom_hline(yintercept=(1/lambda)^2,color='red',size=2) +
ggtitle("Variance of every sample")

```

Above Plot, shows the variances of all the thousand samples and red line is the theoretical variance of the distribution
```{r}
print(paste0("The True Variance is ",true.variance))
print(paste0("The Mean of Sample Variances is ",variance.mean))
```

```{r,fig.cap='Density Plot for Variance of the samples',fig.align='center',fig.height=3.7}
plot.data <- data.frame(x=variance)
g <- ggplot(plot.data, aes(x = x)) + geom_histogram(binwidth=3, 
  colour = "black", aes(y = ..density..)) 
g <- g + geom_vline(xintercept=(1/lambda)^2,color='red',size=2) + geom_vline(xintercept=variance.mean,color='blue',size=2) + 
ggtitle("Distribution of Sample Variance") +
theme(plot.title = element_text(hjust = 0.5))
g
```
\newpage
Blue Line Represents the Mean of Sample Variances and Red Line represents the True Variance of the distribution.
As for mean sample means was estimating true mean of distribution for variance the sample variance will estimate the true variance of distribution.
So again the mean of sample variances is almost equal to true variance of the distribution so Central Limit Theorem is again correct in this case as well

## Comparing Normal Distribution and Sampling Distribution of Mean 
```{r,,fig.align='center',fig.cap="Desnity Plot for Exponential Distribution",fig.height=2.98}

exp.dist <- rexp(nsim,rate=lambda)
exp.dist.data <- data.frame(x=exp.dist)
ggplot(exp.dist.data, aes(x = x)) + geom_histogram(binwidth=2, colour = "black", 
aes(y = ..density..)) + ggtitle("Exponential Distribution") + 
theme(plot.title = element_text(hjust = 0.5))
```


\newpage
```{r,fig.align='center',fig.cap='Distribution of Sample Means',fig.height=3}
g <- ggplot(data, aes(x = x)) + geom_histogram(binwidth=.3, colour = "black",
aes(y = ..density..)) 
g <- g + stat_function(fun=dnorm,args=list(mean=5,sd=5/(nsamples)^0.5),size=2)+ 
ggtitle("Comparing Normal Distribution to Distribution of sample means") + 
theme(plot.title = element_text(hjust = 0.5))
g
```
\newpage
The Distribution of Sample Means is obtained by sampling samples of size = 40 from the exponential distribution and taking averages of them.Central Limit Theorem states that the distribution of sample mean is normal with mean = population mean and Standard Deviation = Standard Error of Mean. 



The above plot compares the normal distribution and sample mean distribution and it clearly visible that sample mean distribution is approximately normal,also if we increase the sample size then it will fit even better.
